---
title: "module2"
author: "ZX"
date: "December 16, 2015"
output: html_document
---

#Q1: What percentage of variation is explained by the 1st principal component in the data set if you:
#a. Do no transformations?
#b. log2(data + 1) transform?
#c. log2(data + 1) transform and subtract row means?
```{r}
#load library
library(devtools)
library(Biobase)
#load data
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/montpick_eset.RData")
load(file=con)
close(con)
mp = montpick.eset
pdata=pData(mp)
edata=as.data.frame(exprs(mp))
fdata = fData(mp)
head(pdata)
head(edata)
head(fdata)
#calculate signatular
svd1 = svd(edata)
names(svd1)
svd1$d[1]^2/sum(svd1$d^2)
plot(svd1$d)
plot(svd1$d^2/sum(svd1$d^2))
#transform data
edata_log = log2(edata+1)
svd1 = svd(edata_log)
svd1$d[1]^2/sum(svd1$d^2)
plot(svd1$d)
plot(svd1$d^2/sum(svd1$d^2))
#transform and substract row mean
edata_centered = edata_log - rowMeans(edata_log)
svd1 = svd(edata_centered)
svd1$d[1]^2/sum(svd1$d^2)
plot(svd1$d)
plot(svd1$d^2/sum(svd1$d^2))
```

#Q2:What is the correlation between the first singular vector and the sample clustering indicator?
```{r}
#load data
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/montpick_eset.RData")
load(file=con)
close(con)
mp = montpick.eset
pdata=pData(mp)
edata=as.data.frame(exprs(mp))
fdata = fData(mp)

#data transform
edata = log2(edata + 1)
edata_centered = edata - rowMeans(edata)
#kmeans cluster
set.seed(333)
kmeans1 = kmeans(edata_centered, centers = 2)
names(kmeans1)
#singular vector
svd1 = svd(edata_centered)
names(svd1)
length(svd1$u[,1])
length(kmeans1$cluster)
plot(kmeans1$cluster,svd1$u[,1])
cor(kmeans1$cluster, svd1$u[,1]) #the result does not match the correct option
```

#Q3:Plot the data for this gene versus the covariate. Can you think of why this model might not fit well?

```{r}
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/bodymap_eset.RData")
load(file=con)
close(con)
bm = bodymap.eset
edata = exprs(bm)
pdata_bm=pData(bm)
edata[1,]
head(pdata_bm)
#fit linear model
lm1 = lm(edata[1,] ~ as.factor(pdata_bm$num.tech.reps))
#plot 
par(mfrow=c(1,2))
plot(edata[1,], pdata_bm$num.tech.reps, col= "blue", pch = 19)
plot(edata[1,], lm1$fitted.values, col= "red", pch = 19)
```

#Q4:Fit a linear model relating he first gene’s counts to the age of the person and the sex of the samples. What is the value and interpretation of the coefficient for age?
```{r}
#load data
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/bodymap_eset.RData")
load(file=con)
close(con)
bm = bodymap.eset
edata = exprs(bm)
pdata_bm=pData(bm)
head(pdata_bm)
#fit linear model to age and sex
lm1 = lm(edata[1,] ~ pdata_bm$age + pdata_bm$gender)
#get coefficients
lm1$coefficients
```

#Q5:What is the dimension of the residual matrix, the effects matrix and the coefficients matrix?
```{r}
#load data
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/montpick_eset.RData")
load(file=con)
close(con)
mp = montpick.eset
pdata=pData(mp)
edata=as.data.frame(exprs(mp))
fdata = fData(mp)
head(pdata)
#transform data
edata = log2(edata + 1)
#fit model for all data
mod = model.matrix(~pdata$population)
fit = lm.fit(mod, t(edata))
names(fit)
head(fit$coefficients)
#check dimensions of coefficients, residuals and effects
dim(fit$coefficients)
dim(fit$residuals)
dim(fit$effects)
```

#Q6:Do this using the lm.fit function (hint: don't forget the intercept). What is the effects matrix?
```{r}
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/montpick_eset.RData")
load(file=con)
close(con)
mp = montpick.eset
pdata=pData(mp)
edata=as.data.frame(exprs(mp))
fdata = fData(mp)
#transform data
edata = log2(edata + 1)
#fit model for all data
mod = model.matrix(~pdata$population)
fit = lm.fit(mod, t(edata))
#effects matrix
names(fit)
fit$effects[1,]
```

#Q7:What is the coefficient for age for the 1,000th gene? Make a plot of the data and fitted values for this gene. Does the model fit well?
```{r}
#load data
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/bodymap_eset.RData")
load(file=con)
close(con)
bm = bodymap.eset
edata = exprs(bm)
pdata_bm=pData(bm)
edata = as.data.frame(edata)
drops = c("ERS025084","ERS025087", "ERS025093")
edata = edata[, !names(edata) %in% drops]
#fit limma model
library(limma)
mod = model.matrix(~pdata_bm$age)
fit_limma = lmFit(edata, mod)
names(fit_limma)
#coefficient of 1000th gene
fit_limma$coefficients[1000,]

mod = model.matrix(~pdata_bm$age + pdata_bm$tissue.type)
fit = lmFit(edata, mod)
```
